---
title: "Useful R packages for the analysis of randomized controlled trials"
author: 
  - "Javier Mancilla Galindo"
date: today
execute: 
  echo: true
  warning: false
toc: false
format:
  docx:
    highlight-style: github
csl: geriatrics.csl
editor: source
---

A full list of useful R packages for the design, monitoring, and analysis of randomized controlled trials can be consulted in this website (CITATION). All these packages can be installed through the `ctv` package by running `ctv::install.views("ClinicalTrials", coreOnly = TRUE)`. 

Citation: 
Ed Zhang, W. G. Zhang, R. G. Zhang (2021). CRAN Task View: Clinical Trial Design, Monitoring, and Analysis. Version 2021-12-29. URL https://CRAN.R-project.org/view=ClinicalTrials.

We would like to make special mention of the following  R packages:  
```{r}

if (!require("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

pacman::p_load(
  table1, # Create table of baseline characteristics
  nlme, # Mixed effects models 
  rms, # Regression modelling strategies.
  Hmisc, # Miscelaneous functions on top of rms
  mice, # Data imputation
  geeCRT # bias-corrected generalized estimating equations (GGE) for 
          # cluster randomized controlled trials
)

```

```{r}
#| include: false
session <- sessionInfo()
# remove clutter
session$BLAS <- NULL
session$LAPACK <- NULL
session$loadedOnly <- NULL

session
```

```{r}
#| include: false
report::cite_packages(session)
```

# Package references 

  - Harrell Jr F (2024). _Hmisc: Harrell Miscellaneous_. R package version 5.1-2, <https://CRAN.R-project.org/package=Hmisc>.
  - Harrell Jr FE (2024). _rms: Regression Modeling Strategies_. R package version 6.8-0, <https://CRAN.R-project.org/package=rms>.
  - Pinheiro J, Bates D, R Core Team (2023). _nlme: Linear and Nonlinear Mixed Effects Models_. R package version 3.1-164, <https://CRAN.R-project.org/package=nlme>. Pinheiro JC, Bates DM (2000). _Mixed-Effects Models in S and S-PLUS_. Springer, New York. doi:10.1007/b98882 <https://doi.org/10.1007/b98882>.
  - R Core Team (2024). _R: A Language and Environment for Statistical Computing_. R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>.
  - Rich B (2023). _table1: Tables of Descriptive Statistics in HTML_. R package version 1.4.3, <https://CRAN.R-project.org/package=table1>.
  - Rinker TW, Kurkiewicz D (2018). _pacman: Package Management for R_. version 0.5.0, <http://github.com/trinker/pacman>.
  - van Buuren S, Groothuis-Oudshoorn K (2011). “mice: Multivariate Imputation by Chained Equations in R.” _Journal of Statistical Software_, *45*(3), 1-67. doi:10.18637/jss.v045.i03 <https://doi.org/10.18637/jss.v045.i03>.
  - Yu H, Li F, Rathouz P, Turner E, Preisser J (2024). _geeCRT: Bias-Corrected GEE for Cluster Randomized Trials_. R package version 1.1.3, <https://CRAN.R-project.org/package=geeCRT>.